### 2020-07-21 data/2020.05.28-TRY/AusTraits.Rmd

TRY

```{r}
library(tidyverse)

bind_rows(
read_tsv("2020.05.28-TRY/tde202052811159.txt", skip=3),
read_tsv("2020.05.28-TRY/tde2020528113232.txt", skip=3)
) %>% distinct() -> TRY
```

```{r}
TRY %>% select(AccSpeciesName, Trait) %>% distinct()
TRY %>% select(AccSpeciesName) %>% distinct()
TRY %>% select(Lat, Lon) %>% distinct()
```

Note, tonnes of repeated values, especially for categorical traits so total number trait values is misleading

```{r}
TRY %>% group_by(AccSpeciesName, Trait) %>% summarise(n=n()) %>% arrange(desc(n))
```

TRY - data for 3820 species, 46,328 trait x species combinations, 1448 locations

AUSTRAITS

```{r}
readRDS("austraits.rds") -> Austraits
Austraits$traits %>% select(species_name, trait_name) %>% distinct()
Austraits$traits$species_name %>% unique()
Austraits$sites
```

Austraits -- data over 650000 records, for 339 traits, for 23,480 species. 282000 trait x species combinations, 22,000 locations